%! Author = deandidion
%! Date = 11.07.25


% Document
\chapter{Interview Guide}

\section*{Interview Guide — AI-Driven Impact Measurement in Public Sector Innovation}

\textbf{Introduction:}

Thank you for agreeing to participate. This interview is part of my thesis research exploring how AI can support measuring impact in public sector innovation. Your insights will help shape a framework for practical and ethical AI applications in this area. The interview will take about 30--45 minutes and your responses will be kept confidential. Please feel free to be candid and share examples.

\vspace{1em}

\section*{Section 1: Background and Role}

\begin{enumerate}
    \item Could you please briefly describe your role and involvement with public sector innovation projects or platforms like the Public Value Academy?
    \item How familiar are you with impact measurement processes in your work? What approaches or tools do you currently use?
\end{enumerate}

\vspace{1em}

\section*{Section 2: Understanding Impact and Measurement}

\begin{enumerate}
    \item What does “impact” mean in the context of public sector innovation, from your perspective?
    \item What are the biggest challenges you face when trying to measure or demonstrate impact?
    \item How important is aligning impact measurement with public values and stakeholder expectations?
\end{enumerate}

\vspace{1em}

\section*{Section 3: AI and Technology in Impact Measurement}

\begin{enumerate}
    \item Have you encountered or used AI or data-driven tools to assist in impact measurement or management? If yes, can you describe your experience?
    \item What potential do you see for AI to improve the transparency, accuracy, or usefulness of impact assessments?
    \item What concerns or risks do you associate with using AI in this context? (e.g., bias, transparency, ethical issues)
\end{enumerate}

\vspace{1em}

\section*{Section 4: Practical and Organizational Considerations}

\begin{enumerate}
    \item What kind of data is typically available or missing that would help in AI-driven impact measurement?
    \item How do you see the role of stakeholders (public officials, citizens, NGOs) in shaping AI-supported impact tools?
    \item What barriers might exist to adopting AI-based solutions in your projects or organization? (e.g., skills, trust, costs)
\end{enumerate}

\vspace{1em}

\section*{Section 5: Vision and Recommendations}

\begin{enumerate}
    \item If you could design the ideal AI tool for impact measurement in public innovation, what features or capabilities would it have?
    \item How can developers ensure that AI supports public values and accountability effectively?
    \item Are there specific ethical guidelines or principles that should govern the use of AI in public sector impact measurement?
\end{enumerate}

\vspace{1em}

\section*{Closing}

\begin{enumerate}
    \item Is there anything else you think is important for understanding how AI could support impact measurement in public sector innovation?
    \item Would you be open to follow-up questions or participating in further discussions during this research?
\end{enumerate}

